Block bootstrap consistency\\under weak assumptions

### 2015-12-31 Paper.tex
[fixup] Make small corrections to the main paper

Our insight is that the expectation of squared blocks of the bootstrap
process can be expressed as a sequence of \emph{contiguous} blocks of
the original process, so the arguments that establish convergence of
the original squared blocks can be applied with only minor changes to
the bootstrapped blocks. Consider the Moving Blocks Bootstrap,%
\footnote{To make this presentation as simple as possible, assume for
  now that $n = M_n J_n$ exactly.} %
for example, and let
\begin{equation*}
  Z_{nj}^* = \tfrac{1}{\sqrt{n}} \sum_{t=(j-1) M_n + 1}^{j M_n} (X_{nt}^* - \bar X_n).
\end{equation*}
Then, conditional on the data, the $Z_{nj}^{*2}$ are independent can
be expected to obey an \lln, so %
$\sum_j (Z_{nj}^{*2} - \E^* Z_{nj}^{*2}) \to^p 0$ %
and the CLT for the bootstrapped array requires $\sum_j \E^*
Z_{nj}^{*2}$ to converge to a positive and finite limit.  But, since
$\E^*$ only averages over the starting point of each block, we have
\begin{align*}
  \E^* Z_{nj}^{*2} &=
  \tfrac{1}{n} \sum_{\tau = 0}^{n-M_n}
  \Big(\tfrac{1}{\sqrt{n}}
  \sum_{t= \tau + 1}^{\tau + M_n} (X_{nt} - \bar X_n) \Big)^2 \\
  &= \tfrac{1}{n} \sum_{\tau_0 = 0}^{M_n-1}
  \sum_{j=1}^{J_n-1}
  \Big(\tfrac{1}{\sqrt{n}}
  \sum_{t= (j - 1) M_n + \tau_0 + 1}^{j M_n + \tau_0} (X_{nt} - \bar X_n) \Big)^2
\end{align*}
after grouping blocks separated by $M_n$ periods. For each $\tau_0$,
the summation can be expected to converge in probability through the
same arguments that were used to establish the \clt\ for the original
array.%
\footnote{Some of the details of the argument will typically need to
  change because the original \clt\ only requires convergence for
  $\tau_0 = 0$, but these details are often incidental to the original
  argument.}%
\footnote{Lemmas~\ref{bootstrap-variance-convergence-lemma} and
  \ref{bootstrapped-zsq-uniform-integrability-lemma} are particularly
  strong demonstrations of this argument.} %
A similar representation is available for the Circular and Stationary
bootstraps.

@@

  To show that $\sigma_n^{*2}$ converges, we can write
  \begin{multline*}
    \sigma_n^{*2} - \sigma^2 = \E^*\Big\{ \sum_{j=1}^{J_n} (Z_{nj}^{*2} - Z_{nj}^{\prime*2}) \Big\}
    + \tfrac{1}{n} \sum_{\tau=0}^{n-1} \E^* \Big\{\sum_{j=1}^{J_n}
    \big( Z_{n}^{\prime}(\tau,M_{nj})^2 - \Em Z_{n}^{\prime}(\tau,M_{nj})^2 \big)\Big\} \\
    + \tfrac{1}{n} \sum_{\tau=0}^{n-1} \E^* \Big\{\sum_{j=1}^{J_n}
    \Em Z_{n}^{\prime}(\tau,M_{nj})^2  - \sigma^2\Big\}.
  \end{multline*}
  Uniform integrability ensures that the convergence in~(\ref{eq:25})
  holds in $L_1$ as well and
  Lemma~\ref{equivalent-convergence-from-lie-lemma} then implies that
  the first term converges to zero in probability.
  Lemma~\ref{lln-for-real-zsq} proves that the second and third
  summation converge to zero in probability.

  Next,
  \begin{equation*}
    \hat \sigma_n^{*2} - \sigma_n^{*2}
    = \sum_{j=1}^{J_n} \big( Z_{nj}^* + (M_{nj}/\sqrt{n}) (\bar X_n - \bar X_{n}^*) \big)
    - \E^*  \sum_{j=1}^{J_n} Z_{nj}^{*2}
  \end{equation*}
  so, in light of the previous arguments, $\hat\sigma_n^{*2} \to^p \sigma^2$ if
  \begin{equation}
    (\bar X_n - \bar X_n^*)^2 \sum_{j=1}^{J_n} M_{nj}^2/n \to^p 0,
  \end{equation}
  which holds by Lemma~\ref{block-length-convergence-lemma} and
  assumption.
\end{proof}

>>>>

Our insight is that the expectation of squared blocks of the bootstrap
process can be expressed as a sequence of \emph{contiguous} blocks of
the original process, so the arguments that establish convergence of
the original squared blocks can be applied with only minor changes to
the bootstrapped blocks. Consider the Moving Blocks Bootstrap,%
\footnote{To make this presentation as simple as possible, assume for
  now that $n = M_n J_n$ exactly.} %
for example, and let
\begin{equation*}
  Z_{nj}^* = \tfrac{1}{\sqrt{n}} \sum_{t=(j-1) M_n + 1}^{j M_n} (X_{nt}^* - \bar X_n).
\end{equation*}
Then, conditional on the data, the $Z_{nj}^{*2}$ are independent can
be expected to obey an \lln, so %
$\sum_j (Z_{nj}^{*2} - \E^* Z_{nj}^{*2}) \to^p 0$ %
and the CLT for the bootstrapped array requires $\sum_j \E^*
Z_{nj}^{*2}$ to converge to a positive and finite limit.  But, since
$\E^*$ only averages over the starting point of each block, we have
\begin{align*}
  \E^* Z_{nj}^{*2} &=
  \tfrac{1}{n} \sum_{\tau = 0}^{n-M_n}
  \Big(\tfrac{1}{\sqrt{n}}
  \sum_{t= \tau + 1}^{\tau + M_n} (X_{nt} - \bar X_n) \Big)^2 \\
  &= \tfrac{1}{n} \sum_{\tau_0 = 0}^{M_n-1}
  \sum_{j=1}^{J_n}
  \Big(\tfrac{1}{\sqrt{n}}
  \sum_{t= (j - 1) M_n + \tau_0 + 1}^{j M_n + \tau_0} (X_{nt} - \bar X_n) \Big)^2
\end{align*}
after grouping blocks separated by $M_n$ periods. For each $\tau_0$,
the summation can be expected to converge in probability through the
same arguments that were used to establish the \clt\ for the original
array.%
\footnote{Some of the details of the argument will typically need to
  change because the original \clt\ only requires convergence for
  $\tau_0 = 0$, but these details are often incidental to the original
  argument.}%
\footnote{Lemmas~\ref{bootstrap-variance-convergence-lemma} and
  \ref{bootstrapped-zsq-uniform-integrability-lemma} are particularly
  strong demonstrations of this argument.} %
A similar representation is available for the Circular and Stationary
bootstraps.

@@

  To show that $\sigma_n^{*2}$ converges, we can write
  \begin{multline}\label{eq:38}
    \sigma_n^{*2} - \sigma^2 = \E^*\Big\{ \sum_{j=1}^{J_n} (Z_{nj}^{*2} - Z_{nj}^{\prime*2}) \Big\}
    + \tfrac{1}{n} \sum_{\tau=0}^{n-1} \E^* \Big\{\sum_{j=1}^{J_n}
    \big( Z_{n}^{\prime}(\tau,M_{nj})^2 - \Em Z_{n}^{\prime}(\tau,M_{nj})^2 \big)\Big\} \\
    + \tfrac{1}{n} \sum_{\tau=0}^{n-1} \E^* \Big\{\sum_{j=1}^{J_n}
    \Em Z_{n}^{\prime}(\tau,M_{nj})^2  - \sigma^2\Big\}.
  \end{multline}
  Uniform integrability ensures that the convergence in~(\ref{eq:25})
  holds in $L_1$ as well and
  Lemma~\ref{equivalent-convergence-from-lie-lemma} then implies that
  the first term in \eqref{eq:38} converges to zero in probability.
  Lemma~\ref{lln-for-real-zsq} proves that the second and third
  summation converge to zero in probability.

  Next,
  \begin{equation*}
    \hat \sigma_n^{*2} - \sigma_n^{*2}
    = \sum_{j=1}^{J_n} \big( Z_{nj}^* + (M_{nj}/\sqrt{n}) (\bar X_n - \bar X_{n}^*) \big)^2
    - \E^*  \sum_{j=1}^{J_n} Z_{nj}^{*2}
  \end{equation*}
  so, in light of the previous arguments, $\hat\sigma_n^{*2} \to^p \sigma^2$ if
  \begin{equation}
    (\bar X_n - \bar X_n^*)^2 \sum_{j=1}^{J_n} M_{nj}^2/n \to^p 0,
  \end{equation}
  which holds by Lemma~\ref{block-length-convergence-lemma} and
  assumption.
\end{proof}
